\emph{MixFishSim}: highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2019-06-27 write_up/F_changes.tex
Revised text

% latex table generated in R 3.5.2 by xtable 1.8-3 package
% Mon Jan 28 16:57:19 2019
\begin{longtable}{rllrrrlllr}
\caption{Fishing mortality effects of the closure scenarios} \\ 
  \hline
scenario & metric & pop & before & after & diff (\%) & timescale & basis & data\_type & resolution \\ 
  \hline
  9 & F & spp\_3 & 1.08 & 0.29 & -73.47 & weekly & high\_pop & real\_pop & 1.00 \\ 
   10 & F & spp\_3 & 1.08 & 0.29 & -72.94 & monthly & high\_pop & real\_pop & 1.00 \\ 
   11 & F & spp\_3 & 1.08 & 0.35 & -68.04 & yearly & high\_pop & real\_pop & 1.00 \\ 
   45 & F & spp\_3 & 1.08 & 0.58 & -46.70 & yearly & high\_pop & commercial & 20.00 \\ 
    1 & F & spp\_3 & 1.08 & 0.58 & -46.21 & weekly & high\_pop & commercial & 1.00 \\ 
   23 & F & spp\_3 & 1.08 & 0.59 & -45.27 & weekly & high\_pop & real\_pop & 5.00 \\ 
    2 & F & spp\_3 & 1.08 & 0.59 & -45.06 & monthly & high\_pop & commercial & 1.00 \\ 
    7 & F & spp\_3 & 1.08 & 0.60 & -44.48 & yearly & high\_pop & survey & 1.00 \\ 
   24 & F & spp\_3 & 1.08 & 0.61 & -43.20 & monthly & high\_pop & real\_pop & 5.00 \\ 
    3 & F & spp\_3 & 1.08 & 0.64 & -40.82 & yearly & high\_pop & commercial & 1.00 \\ 
   25 & F & spp\_3 & 1.08 & 0.65 & -39.94 & yearly & high\_pop & real\_pop & 5.00 \\ 
   17 & F & spp\_3 & 1.08 & 0.67 & -38.11 & yearly & high\_pop & commercial & 5.00 \\ 
   15 & F & spp\_3 & 1.08 & 0.71 & -34.38 & weekly & high\_pop & commercial & 5.00 \\ 
   43 & F & spp\_3 & 1.08 & 0.71 & -34.31 & weekly & high\_pop & commercial & 20.00 \\ 
   16 & F & spp\_3 & 1.08 & 0.73 & -32.58 & monthly & high\_pop & commercial & 5.00 \\ 
   51 & F & spp\_3 & 1.08 & 0.78 & -27.92 & weekly & high\_pop & real\_pop & 20.00 \\ 
   37 & F & spp\_3 & 1.08 & 0.78 & -27.76 & weekly & high\_pop & real\_pop & 10.00 \\ 
   39 & F & spp\_3 & 1.08 & 0.79 & -26.98 & yearly & high\_pop & real\_pop & 10.00 \\ 
   38 & F & spp\_3 & 1.08 & 0.81 & -25.47 & monthly & high\_pop & real\_pop & 10.00 \\ 
   21 & F & spp\_3 & 1.08 & 0.81 & -25.21 & yearly & high\_pop & survey & 5.00 \\ 
   35 & F & spp\_3 & 1.08 & 0.81 & -25.05 & yearly & high\_pop & survey & 10.00 \\ 
   44 & F & spp\_3 & 1.08 & 0.87 & -19.91 & monthly & high\_pop & commercial & 20.00 \\ 
   52 & F & spp\_3 & 1.08 & 0.88 & -18.39 & monthly & high\_pop & real\_pop & 20.00 \\ 
   30 & F & spp\_3 & 1.08 & 0.96 & -11.06 & monthly & high\_pop & commercial & 10.00 \\ 
   29 & F & spp\_3 & 1.08 & 0.98 & -9.80 & weekly & high\_pop & commercial & 10.00 \\ 
   31 & F & spp\_3 & 1.08 & 1.03 & -4.36 & yearly & high\_pop & commercial & 10.00 \\ 
   53 & F & spp\_3 & 1.08 & 1.06 & -1.64 & yearly & high\_pop & real\_pop & 20.00 \\ 
   49 & F & spp\_3 & 1.08 & 1.07 & -1.01 & yearly & high\_pop & survey & 20.00 \\ 
   \hline
\hline
\label{tab:6}
\end{longtable}

>>>>

% latex table generated in R 3.6.0 by xtable 1.8-4 package
% Thu Jun 27 18:22:53 2019
\begin{longtable}{rllrrrlllr}
\caption{Fishing mortality effects of the closure scenarios. Results show the
	fishing mortality before the closure (f\_before) and after the closure
	(f\_after) and the percentage change in f (f\_change). The results are
	ordered by most effective scenario first, least effective last.)} \\ 
  \hline
scenario & metric & pop & f\_before & f\_after & f\_change & timescale & basis & data\_type & resolution \\ 
  \hline
  9 & F & spp\_3 & 1.08 & 0.29 & -73.47 & weekly & high\_pop & real\_pop & 1.00 \\ 
   10 & F & spp\_3 & 1.08 & 0.29 & -72.94 & monthly & high\_pop & real\_pop & 1.00 \\ 
   11 & F & spp\_3 & 1.08 & 0.35 & -68.04 & yearly & high\_pop & real\_pop & 1.00 \\ 
   45 & F & spp\_3 & 1.08 & 0.58 & -46.70 & yearly & high\_pop & commercial & 20.00 \\ 
    1 & F & spp\_3 & 1.08 & 0.58 & -46.21 & weekly & high\_pop & commercial & 1.00 \\ 
   23 & F & spp\_3 & 1.08 & 0.59 & -45.27 & weekly & high\_pop & real\_pop & 5.00 \\ 
    2 & F & spp\_3 & 1.08 & 0.59 & -45.06 & monthly & high\_pop & commercial & 1.00 \\ 
    7 & F & spp\_3 & 1.08 & 0.60 & -44.48 & yearly & high\_pop & survey & 1.00 \\ 
   24 & F & spp\_3 & 1.08 & 0.61 & -43.20 & monthly & high\_pop & real\_pop & 5.00 \\ 
    3 & F & spp\_3 & 1.08 & 0.64 & -40.82 & yearly & high\_pop & commercial & 1.00 \\ 
   25 & F & spp\_3 & 1.08 & 0.65 & -39.94 & yearly & high\_pop & real\_pop & 5.00 \\ 
   17 & F & spp\_3 & 1.08 & 0.67 & -38.11 & yearly & high\_pop & commercial & 5.00 \\ 
   15 & F & spp\_3 & 1.08 & 0.71 & -34.38 & weekly & high\_pop & commercial & 5.00 \\ 
   43 & F & spp\_3 & 1.08 & 0.71 & -34.31 & weekly & high\_pop & commercial & 20.00 \\ 
   16 & F & spp\_3 & 1.08 & 0.73 & -32.58 & monthly & high\_pop & commercial & 5.00 \\ 
   51 & F & spp\_3 & 1.08 & 0.78 & -27.92 & weekly & high\_pop & real\_pop & 20.00 \\ 
   37 & F & spp\_3 & 1.08 & 0.78 & -27.76 & weekly & high\_pop & real\_pop & 10.00 \\ 
   39 & F & spp\_3 & 1.08 & 0.79 & -26.98 & yearly & high\_pop & real\_pop & 10.00 \\ 
   38 & F & spp\_3 & 1.08 & 0.81 & -25.47 & monthly & high\_pop & real\_pop & 10.00 \\ 
   21 & F & spp\_3 & 1.08 & 0.81 & -25.21 & yearly & high\_pop & survey & 5.00 \\ 
   35 & F & spp\_3 & 1.08 & 0.81 & -25.05 & yearly & high\_pop & survey & 10.00 \\ 
   44 & F & spp\_3 & 1.08 & 0.87 & -19.91 & monthly & high\_pop & commercial & 20.00 \\ 
   52 & F & spp\_3 & 1.08 & 0.88 & -18.39 & monthly & high\_pop & real\_pop & 20.00 \\ 
   30 & F & spp\_3 & 1.08 & 0.96 & -11.06 & monthly & high\_pop & commercial & 10.00 \\ 
   29 & F & spp\_3 & 1.08 & 0.98 & -9.80 & weekly & high\_pop & commercial & 10.00 \\ 
   31 & F & spp\_3 & 1.08 & 1.03 & -4.36 & yearly & high\_pop & commercial & 10.00 \\ 
   53 & F & spp\_3 & 1.08 & 1.06 & -1.64 & yearly & high\_pop & real\_pop & 20.00 \\ 
   49 & F & spp\_3 & 1.08 & 1.07 & -1.01 & yearly & high\_pop & survey & 20.00 \\ 
   \hline
\hline
\label{tab:6}
\end{longtable}

### 2019-06-27 write_up/mixfishsim_ecolmod.tex
Revised text

\begin{enumerate}
	\item How does sampling-derived data reflects the underlying population
		structures? 	
	\item How does data aggregation and source impact on monitoring spatial
		fisheries management measures ? \end{enumerate}

@@

It can be seen from a single vessels movements during a trip that the vessel
exploits four different fishing grounds, three of them multiple time (Figure
\ref{S-fig:11}), while across multiple trips fishing grounds that are further
apart are fished (Figure \ref{S-fig:12}). These different locations relate to
areas where the highest revenue were experienced, as shown by
\added[id=PD]{Figure} \ref{S-fig:13}, which shows several trips for the vessel
overlaid on the value field (sum of the population densities $\times$ price).
Vessels from the same fleet (and therefore targeting preference) exploit
similar but slightly different fishing grounds depending on their own personal
experience during the explore phase of the fishery (\added[id=PD]{Figure}
\ref{S-fig:14}), which is the result of the correlated random walk step
function, with distance moved during the exploitation phase related to the
revenue experienced on the fishing ground (\replaced[id=PD]{Figure}{while}
\ref{S-fig:15}).\todo{\added[id=JJ]{Move some of the
		supplementary figures to the manuscript}}

@@

\subsection{Example research question scenario}

To illustrate the capabilities on \emph{MixFishSim}, we investigate the
influence of the temporal and spatial resolution of different data sources on
the reduction in catches of a population given spatial closures. To do so, we
set up a simulation to run for 50 years based on a 100 $\times$ 100 square grid
(undetermined units), with five fleets of 20 vessels each and four fish
populations. Fishing takes place four times a day per vessel and five days a
week, while population movement is every week.\todo{\added[id=JJ]{move to start
		of methods section}\added[id=PD]{I think ecological modelling
		wants the 'calculations' section here..will check}} \\

@@

The following steps are undertaken to determine closures:
\todo{\added[id=JJ]{Procedure unclear. Refer to symbols in methods section or
		switch order starting with description of data type
		etc..}\added[id=PD]{Yes, will redo}}
\begin{enumerate}
	\item Extract data source
	\item Aggregate according to desired spatial and temporal resolution
	\item Interpolate across entire area at desired resolution
	\item Close area covering top 5 \% of catch 
\end{enumerate}
In total 56 closure scenarios were run which represent combinations of:

\begin{itemize}
	\item \textbf{data types:} commercial logbook data, survey data and
		`real population',
	\item \textbf{temporal resolutions:} weekly, monthly and yearly
		closures,
	\item \textbf{spatial resolutions:} 1 x 1 grid, 5 x 5 grid, 10 x 10
		grid and 20 x 20 grid,
	\item \textbf{closure basis:} high catch rates of protected species, or
		high ratio of protected species v secondary species.
\end{itemize}

@@

In order to answer the question of how sampling-derived data reflects the
underlying population structure we compare different spatial and temporal
aggregations of the simulated population distributions to:
\begin{enumerate}[label=\alph*)]
	\item \textbf{fisheries-independent data:} the inferred population from
		a fixed-site sampling survey design as commonly used for
		fisheries monitoring purposes;
	\item \textbf{fisheries-dependent data:} the inferred population from
		our fleet model which includes fishery-induced sampling
		dynamics.
\end{enumerate}

The consequences of different spatial aggregations of the data are shown in
Figure \ref{fig:1}, which represents the aggregation of catch from each of the
data sources over a ten-year period (to average seasonal patterns) at different
spatial resolutions. \\

The finer spatial grid for the real population (top left) and commercial data
(top middle) show visually similar patterns, though there are large unsampled
areas in the commercial data from a lack of fishing activity (particularly in
the lower left part of the sampling domain). The survey data at this spatial
resolution displays very sparse information about the spatial distributions of the
populations. The slightly aggregated data on a 5 x 5 grid shows similar
patterns and, while losing some of the spatial detail, there remains good
consistency between the `real population' and the commercial data. Survey data
starts to pick out some of the similar patterns as the other data sources, but
lacks coverage. The spatial catch information on a 10 x 10 and 20 x 20 grid
lose a significant amount of information about the spatial resolutions for all
data sources, and some differences between the survey, commercial and 'real
population' data emerge. \\

Figure \ref{fig:2} shows the consequences of different temporal aggregations of
the data over a ten-year period, with weekly (top), monthly (middle) and yearly
(bottom) catch compositions from across an aggregated 20 x 20 area. \\

As can be seen by comparison to the 'real population', the monthly aggregation
captures the major patterns seen in the weekly data, albeit missing more subtle
differences. The yearly data results in a constant catch pattern due to the
aggregation process (sometimes known as an aggregation bias). The commercial
data on a weekly basis shows some of the same patterns as the 'real
population', though the first species (in red) is less well represented and
some weeks are missing catches from the area. The monthly data shows some
consistency between the 'real population' and commercial data for species 2 -
4, though species 1 remains under-represented. On an annual basis,
interestingly the commercial data under represents the first species (in red)
while the survey over represents species 1. This is likely due to the
biases in commercial sampling, with the fisheries not targeting the areas where
species 1 are present, and the biases in the survey sampling from over
representation of the spatial distribution. \\

We implemented a spatial closure using the different data sources and spatial
and temporal aggregations as outlined in the protocol in Section 3.4. We used
this to assess the efficacy of a closure in reducing fishing mortality on
species 1, given availability of data and its use at different resolutions in
order to evaluate the trade-offs in data sources. Figure \ref{fig:3} shows the
trend in fishing mortality for each species simulated (columns) given the data
sources (rows), temporal aggregations (colour lines) and spatial aggregations
(line-styles), while Figure \ref{fig:4} shows the change in fishing mortality
from before the closure (year 29) to after the closure (year 50). \\

For the closures based on 'real population' (bottom row), the most
disaggregated data (a weekly timescale and 1 x 1 resolution) was most
effective, reducing fishing mortality on species 1 (left) by $\sim$ 60 \%. Next
was the monthly closures ($<$ $\sim$ 30 \%).  The least effective were the
yearly closures (blue lines) at all spatial resolutions, which conversely
resulted in increased fishing mortalities ($>$ 30 \%). \\

For the survey data, which can only be implemented on a yearly timescale, the
closures had no effect at any data resolution. The results are identical for
the different data resolutions except 20 x 20, which is why you can't see more
than 2 points. This is because of the sparsity of the sampling locations.\\

For the commercial data, the most effective closure scenario was based on 1 x 1
data at a monthly temporal resolution. This results in $\sim$ 10 \% reduction
in F for species 1. \textcolor{blue}{This was the only closure scenario to have
	positive effect according to Figure \ref{fig:4}, though looking at the
	trend in Figure \ref{fig:3} this looks more related to the continued
	increased in F trend, as other scenarios had an initial effect}.
Interestingly the monthly data scenario was more effective than weekly data,
which we posit is due to the increased data available from the commercial
sampling across a month compared to a week. Commercial data used at an annual
time-step was ineffective in bringing fishing mortality down for species 1. \\

Given the scenarios above, it seems clear that spatial disaggregation is more
important than the temporal disaggregation of the commercial data, except when
its used at an annual time-frame, which is the scenario that gave the worst
results.

For the other species in the simulation (population 2 - 4) there was little
difference in fishing mortalities across scenarios.

\textcolor{blue}{Note: The monthly commercial data scenario is the most
	effective of the realistic scenarios, as the 'real population' can only
	be seen as a baseline comparison.}

@@

@@

\section*{Funding} This work was supported by the MARES doctoral training
program; and the Centre for Environment, Fisheries and Aquaculture Science
seedcorn program.

@@

\clearpage
\footnotesize
\input{F_changes_ratio.tex}

@@

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../analysis/F_trendsREV}
	\caption{Comparison of closure scenarios - Fishing mortality trends.
		Only the scenarios based on high catch rates of population 3
		are shown.}
	\label{fig:3}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width =\linewidth]{../analysis/Overview_plot_highPopRev}
	\caption{Comparison of closure scenarios. Points indicate the
		difference between the fishing mortality pre-closure (year 29)
		and post-closure (year 50) for population 3. Only the scenarios
		based on high catch rates of population 3 are shown.}
	\label{fig:4}
\end{figure}

>>>>

\begin{enumerate}
	\item How does sampling-derived data reflects the underlying population
		structures? 	
	\item How does data aggregation and source impact on spatial fisheries
	management measures?
\end{enumerate}

@@

@@

\subsection{Example research question}

To illustrate the capabilities of \emph{MixFishSim}, we investigate the
influence of the temporal and spatial resolution of different data sources on
the reduction in catches of a population given spatial closures. To do so, we
set up a simulation to run for 50 years based on a 100 $\times$ 100 square grid
(undetermined units), with five fleets of 20 vessels each and four fish
populations. Fishing takes place four times a day per vessel and five days a
week, while population movement is every week.

@@

The following steps are undertaken to determine closures:
\todo{\added[id=JJ]{Procedure unclear. Refer to symbols in methods section or
		switch order starting with description of data type
		etc..}\added[id=PD]{Yes, will redo}}
\begin{enumerate}
	\item Extract data source
	\item Aggregate according to desired spatial and temporal resolution
	\item Interpolate across entire area at desired resolution
	\item Close area covering top 5 \% of catch 
\end{enumerate}
In total 28 closure scenarios were run which represent combinations of:

\begin{itemize}
	\item \textbf{data types:} commercial logbook data, survey data and
		`real population',
	\item \textbf{temporal resolutions:} weekly, monthly and yearly
		closures,
	\item \textbf{spatial resolutions:} 1 x 1 grid, 5 x 5 grid, 10 x 10
		grid and 20 x 20 grid,
	\item \textbf{closure basis:} highest 5 \%  of catch rates for the
		protected species
\end{itemize}

@@

\subsection{Simulation dynamics}

It can be seen from a single vessels movements during a trip that the vessel
exploits four different fishing grounds, three of them multiple times (Figure
\ref{S-fig:11}), while across several trips fishing grounds that are further
apart are fished (Figure \ref{S-fig:12}). These different locations relate to
areas where the highest revenue were experienced, as shown by
\added[id=PD]{Figure} \ref{S-fig:13}, where several trips for the vessel
overlaid on the revenue field, i.e. 
$$\sum^c_{c=1}\sum^s_{s=1} B_{s,c} \cdot Q_{s,c}$$

Vessels from the same fleet (and therefore targeting preference) exploit
similar but slightly different fishing grounds depending on their own personal
experience during the explore phase of the fishery (\added[id=PD]{Figure}
\ref{S-fig:14}), which is the result of the randomness in the correlated random
walk step function, with distance moved during the exploitation phase and the
direction stochastically related to the revenue experienced on the fishing
ground (\replaced[id=PD]{Figure}{while}
\ref{S-fig:15}).\todo{\added[id=JJ]{Move some of the supplementary figures to
		the manuscript}}

\subsection{How does sampling-derived data reflect the underlying population
	structure?}

In order to answer this question we compare different spatial and temporal
aggregations of the simulated population distributions to:
\begin{enumerate}[label=\alph*)]
	\item \textbf{fisheries-independent data:} the inferred population from
		a fixed-site sampling survey design as commonly used for
		fisheries monitoring purposes;
	\item \textbf{fisheries-dependent data:} the inferred population from
		our fleet model which includes fishery-induced sampling
		dynamics.
\end{enumerate}

Figure \ref{fig:1} shows the aggregated catch composition from each of the data
sources over a ten-year period (to average seasonal patterns) at different
spatial resolutions. The finer spatial grid for the real population (top left)
and commercial data (top middle) show visually similar patterns, though there
are large unsampled areas in the commercial data from a lack of fishing
activity (particularly in the lower left part of the sampling domain). The
survey data at this spatial resolution displays very sparse information about
the spatial distributions of the populations. The slightly aggregated data on a
5 x 5 grid shows similar patterns and, while losing some of the spatial detail,
there remains good consistency between the `real population' and the commercial
data. Survey data starts to pick out some of the similar patterns as the other
data sources, but lacks coverage. The spatial catch information on a 10 x 10
and 20 x 20 grid lose a significant amount of information about the spatial
resolutions for all data sources, and some differences between the survey,
commercial and 'real population' data emerge. \\

Figure \ref{fig:2} shows the consequences of different temporal aggregations of
the data over a ten-year period, with weekly (top), monthly (middle) and yearly
(bottom) catch compositions from across an aggregated 20 x 20 area. As can be
seen by comparison to the 'real population', the monthly aggregation captures
the major patterns seen in the weekly data, albeit missing more subtle
differences. The yearly data results in a constant catch pattern due to the
aggregation process (sometimes known as an aggregation bias). The commercial
data on a weekly basis shows some of the same patterns as the 'real
population', though the first species (in red) is less well represented and
some weeks are missing catches from the area. The monthly data shows some
consistency between the 'real population' and commercial data for species 2 -
4, though species 1 remains under-represented. On an annual basis,
interestingly the commercial data under represents the first species (in red)
while the survey over represents species 1. This is likely due to the biases in
commercial sampling, with the fisheries not targeting the areas where species 1
are present, and the biases in the survey sampling from over representation of
the spatial distribution.

\subsection{How does data aggregation and source impact on spatial fisheries
	management measures?}

We implemented a spatial closure using the different data sources and spatial
and temporal aggregations as outlined in the protocol in Section 3.4. We used
this to assess the efficacy of a closure in reducing fishing mortality on
species 3, given availability of data and its use at different resolutions in
order to evaluate the trade-offs in data sources.

The trend in fishing mortality for each species show that in most cases the
fishery closure was successful in reducing fishing mortality on the species of
interest (species 3; Figure \ref{fig:3}), though interestingly the largest
reductions in fishing mortality happened immediately after the closures,
following which the fisheries "adapted" to the closures and fishing mortality
increased again somewhat. The exception to the success was the closures
implemented based on the coarsest spatial (20 x 20) and temporal resolution
which were ineffective with all data sources. As expected, closures based on
the "known" population distribution were most effective, with differing degrees
of success using the commercial data. Fishing mortality rates on the other
species changed in different proportions, depending on whether the displaced
fishing effort moved to areas where the populations were in more greater or
lesser density. \\

A regression tree highlights that the most factor contributing to differences
in fishing mortality before and after the closure was the population (72 \%
showing that the closures were effective for population 3), followed by data
resolution (21 \%), data type (7 \%) with the least important factor the
timescale ($<$ 1 \%). In general the finer the spatial resolution of the data
used the greater reduction in fishing mortality for population 3 after the
closures (Figure\ref{fig:4}). The notable outliers are the commercial data at
the coarsest spatial resolution (20 x 20) at a yearly and weekly timescale,
where closures were nearly as effective as the fine-scale resolution. In this
case the closures were sufficiently large to protect the population (and was as
effective as when the closure was based on the "real population") but this may
have consequences in terms of restricting a much larger area than necessary. \\

@@

It seems clear that spatial disaggregation is more important than the temporal
disaggregation of the commercial data... WHY

@@

\section*{Funding} This work was supported by the MARES doctoral training
program (MARES\_14\_15) and the Centre for Environment, Fisheries and
Aquaculture Science seedcorn program (DP227AC).

@@

@@

\begin{figure}[!ht]
	\includegraphics[width = \linewidth]{../analysis/F_trendsREV}
	\caption{Comparison of closure scenarios effect on fishing mortality
		trends. Line colour denotes the timescale, while linestyle
		denotes the spatial resolution.}
	\label{fig:3}
\end{figure}

\begin{figure}[!ht]
	\includegraphics[width =\linewidth]{../analysis/f_diff_effectiveness}
	\caption{Comparison of closure scenario effectiveness based on
		different spatial and temporal resolutions.}
	\label{fig:4}
\end{figure}
